File-based localization of numerical perturbations in data analysis pipelines

### 2020-04-06 paper.tex

The HCP developed a set of pre-processing pipelines to process structural,
functional, and diffusion MRI data acquired in the project. We focus on HCP
pre-processing pipelines for structural data, and particularly
on PreFreeSurfer and FreeSurfer. 
A detailed description of the analyses done by these
pipelines is available in~\cite{glasser2013}. 
In summary, PreFreeSurfer consists of the following steps: 
\begin{itemize}
\item Gradient Distortion Correction (DC), 
\item Alignment and Anatomical Average (AAve), T1w(s), T2w(s),  
\item Anterior/Posterior Commissure Alignment (ACPC-A), 
\item Brain Extraction (BExt), 
\item \tristan{Ali, why isn't BBR cross-modal registration here and in the heatmap?}
\item Bias Field Correction (BFC), 
\item Atlas-Registration (AR).
\end{itemize}
And FreeSurfer consists of the following ones:
\begin{itemize}
\item Image downsampling, 
\item T1w image registration, 
\item T1w image segmentation,
\item Surface placement, 
\item Surface registration.
\end{itemize}
The average processing time per subject is approximately 2~hours for
PreFreeSurfer and 8~hours for FreeSurfer. The average output file size is
approximately 2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
%\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}

@@

Figure~\ref{fig:pfs_freq} shows the frequency of non-reproducible pipeline processes
in PreFreeSurfer. 
Differences were observed in linear registration 
with FSL \flirt (in ACPC-Alignment, Brain Extraction, Distortion Correction, and
Atlas Registration), in non-linear registration with FSL \fnirt (in Brain Extraction 
and Atlas Registration), and in image warping with FSL \texttt{new\_invwarp} (in Brain Extraction 
and Atlas Registration). Differences were also observed in image mean 
computations with FSL maths  (in Anatomical Average). 
A complete provenance graph of the PreFreeSurfer is illustrated in Figure~\ref{fig:complete_pfs}
\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{images/pfs-intransient} 
    \caption{A complete provenance graph from the PreFreesurfer pipeline.
     Full-resolution image is available 
     \href{https://drive.google.com/file/d/1xonnMq4hLEe6GwMp4P7umikvrIAeflDe/view?usp=sharing}{here}.}
    \label{fig:complete_pfs}
\end{figure}
% Between subject differences can be a result of different subject types. 
% For instance, the subjects contain two T1 images need more calculations 
% to produce an average image between them.

% make figure in two columns use {figure*}
\begin{figure*}
\centering
  \includegraphics[width=\textwidth]{./data/pfs-20sbj-output/heatmap.png}
  \caption{Heatmap of non-reproducible processes across PreFreeSurfer pipeline steps (N=20). 
  Each cell represents the occurrence of a particular command line in a
  pipeline step among Anatomical Average (AAve), Anterior/Posterior
  Commissure Alignment (ACPC-A), Brain Extraction (BExt), Bias Field
  Correction (BFC), or Atlas-Registration (AR). Cell labels indicate the
  number of subjects where the corresponding process appears. For example,
  the \flirt tool was invoked 6 times in step DC for each of the 20
  subjects: 2 instances weren't reproducible in 19 subjects, 3
  instances were always reproducible, and 1 instance wasn't reproducible in
  17 subjects.
        }
  \label{fig:pfs_freq}
\end{figure*}

>>>>

The HCP developed a set of pre-processing pipelines to process structural,
functional, and diffusion MRI data acquired in the project. We focus on HCP
pre-processing pipelines for structural data, and particularly
on PreFreeSurfer and FreeSurfer. 
A detailed description of the analyses done by these
pipelines is available in~\cite{glasser2013}. 
In summary, PreFreeSurfer consists of the following steps: 
\begin{itemize}
\item Gradient Distortion Correction (DC), 
\item Alignment and Anatomical Average (AAve), T1w(s), T2w(s),  
\item Anterior/Posterior Commissure Alignment (ACPC-A), 
\item Brain Extraction (BExt), 
\item \tristan{Ali, why isn't BBR cross-modal registration here and in the heatmap?}
\item Bias Field Correction (BFC), 
\item Atlas-Registration (AR).
\end{itemize}
And FreeSurfer consists of the following ones:
\begin{itemize}
\item Image downsampling, 
\item T1w image registration, 
\item T1w image segmentation,
\item Surface placement, 
\item Surface registration.
\end{itemize}
The average processing time per subject is approximately 2~hours for
PreFreeSurfer and 8~hours for FreeSurfer. The average output file size is
approximately 2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
%\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
The \reprozip traced files show that approximately PreFreeSurfer opened  
83K files and created 7.7K processes, and FreeSurfer opened 62K files and 
created 4K processes in total for a single subject.

@@

Figure~\ref{fig:pfs_freq} shows the frequency of non-reproducible pipeline processes
in PreFreeSurfer. 
Differences were observed in linear registration 
with FSL \flirt (in ACPC-Alignment, Brain Extraction, Distortion Correction, and
Atlas Registration), in non-linear registration with FSL \fnirt (in Brain Extraction 
and Atlas Registration), and in image warping with FSL \texttt{new\_invwarp} (in Brain Extraction 
and Atlas Registration). Differences were also observed in image mean 
computations with FSL maths  (in Anatomical Average). 
A complete provenance graph of the PreFreeSurfer is illustrated in Figure~\ref{fig:complete_pfs}
\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{images/pfs-labeled.png} 
    \caption{A complete provenance graph from the PreFreesurfer pipeline.
     Full-resolution image is available 
     \href{https://drive.google.com/file/d/1BKq5ZXIVr3bkij88SmeW7ulWOExSCm7n/view?usp=sharing}{here}.}
    \label{fig:complete_pfs}
\end{figure}
% Between subject differences can be a result of different subject types. 
% For instance, the subjects contain two T1 images need more calculations 
% to produce an average image between them.

% make figure in two columns use {figure*}
\begin{figure*}
\centering
  \includegraphics[width=\textwidth]{./data/pfs-20sbj-output/heatmap.pdf}
  \caption{Heatmap of non-reproducible processes across PreFreeSurfer pipeline steps (N=20). 
  Each cell represents the occurrence of a particular command line in a
  pipeline step among Anatomical Average (AAve), Anterior/Posterior
  Commissure Alignment (ACPC-A), Brain Extraction (BExt), Bias Field
  Correction (BFC), or Atlas-Registration (AR). Cell labels indicate the
  number of subjects where the corresponding process appears. For example,
  the \flirt tool was invoked 6 times in step DC for each of the 20
  subjects: 2 instances weren't reproducible in 19 subjects, 3
  instances were always reproducible, and 1 instance wasn't reproducible in
  17 subjects.
        }
  \label{fig:pfs_freq}
\end{figure*}
